Search-based Synthesis of Network Intrusion Detection Rules

### 2019-09-24 paper/paper.tex
pass on technique

These intrusion detection mechanisms are complementary. Rule-based
NIDS are restricted to a fixed set of rules and not all rules are
simulneteously enabled (for efficiency). Attackers constantly create
new attacks, and manually writing rules is tedious and
error-prone. However, the semantics of the specification language of
rules is clear making it easy for system administrator to understand
them. As such, rule-based systems are very popular in industry.
Anomaly-based NIDS are not restricted to a fixed set of rules; they
learn common behaviors and dynamically detect uncommon behavior. They
typically use machine learning algorithm to model event
sequences~\cite{7579764}. Unfortunately, anomaly-based NIDS can raise
false alarms as behavior models are fundamentally imprecise. To sum
up, both approaches have positive and negative aspects.

@@

This paper makes the following contributions. \Fix{...}

@@

The SYN Flood attack exploits a vulnerability in the TCP/IP handshake
to establish a TCP connection~\cite{cloudfare-synflood}. The handshake
works as follows. First, a client sends a SYN packet to the server,
requesting a connection. Second, the server responds with a SYN-ACK
packet to the client. Third, the client responds with an ACK message
and the connection is established. Aware of the protocol, an attacker
sends multiple SYN packets to different ports of a server, often using
fake IP addresses. Then, after the server responds with a SYN-ACK
packet, the client sends another SYN to avoid the connection to time
out. Without proper protection, the server accepts the connection
requests and eventually legitimate connection requests cannot be
satisfied due to resource exhaustion.

@@

\begin{figure*}[t]
\centering
\includegraphics[trim=60 400 65 0,clip,width=0.7\textwidth]{figs/nids-workflow}
\caption{\tname\ workflow.}
\label{fig:overview}
\end{figure*}

@@

The technique synthesizes rules in two steps. First, it produces a
potentially overspecified rule that captures the negative example that
is provided as input. Then, it uses positive examples to eliminate
unnecessary options from the rule. The goal of \tname\ is to create a
rule that only captures the negative traffic.

Figure~\ref{fig:overview} shows the workflow of \tname{} as a pipeline
of two components, each implementing one of the steps mentioned above.
The goal of the first component is to produce a rule that captures the
negative traffic.  First, it initializes the rule encoding with random
values. Only options associated with the used protocol are included in
the encoding. Then, \tname\ optimizes the rule options until \suri\ is
able to capture the attack.  Note that, as all options need to be
satisfied for the traffic to be captured, at this point, any subset of
options would capture the negative traffic. The second component of
the pipeline is initialized with the rule produced in the previous
step. That rule captures the negative traffic but could also capture
benign traffic. This component systematically discards options from
the rule encoding until no positive traffic is captured.

\Fix{describes negative and positive traffic}

>>>>

These intrusion detection mechanisms are complementary. Rule-based
NIDS are restricted to a fixed set of rules and not all rules are
simultaneously enabled (for efficiency). Attackers constantly create
new attacks, and manually writing rules is tedious and
error-prone. However, the semantics of the specification language of
rules is clear making it easy for system administrator to understand
them. As such, rule-based systems are very popular in industry.
Anomaly-based NIDS are not restricted to a fixed set of rules; they
learn common behaviors and dynamically detect uncommon behavior. They
typically use machine learning algorithm to model event
sequences~\cite{7579764}. Unfortunately, anomaly-based NIDS can raise
false alarms as behavior models are fundamentally imprecise. To sum
up, both approaches have positive and negative aspects.

@@

This paper makes the following contributions.

@@

The SYN Flood attack exploits a vulnerability in the TCP/IP handshake
to establish a TCP connection~\cite{cloudfare-synflood}. The handshake
works as follows in normal circumstances. First, a client sends a SYN
packet to the server, requesting a connection. Second, the server
responds with a SYN-ACK packet to the client. Third, the client
responds with an ACK message and the connection is established. Aware
of the protocol, an attacker sends multiple SYN packets to different
ports of a server, often using fake IP addresses. Then, after the
server responds with a SYN-ACK packet, the client keeps sending other
SYN to avoid the connection to time out. Without proper protection,
the server accepts these malicious requests and eventually legitimate
requests cannot be satisfied due to resource exhaustion.

@@

\begin{figure*}[t]
\centering
\includegraphics[trim=10 360 182 130,clip,width=0.7\textwidth]{figs/nids-workflow}
\caption{\tname\ workflow.}
\label{fig:overview}
\end{figure*}

@@

The technique synthesizes rules in two steps. First, it produces a
potentially overspecified rule that captures the negative example
provided on input. Then, it uses positive examples to eliminate
unnecessary options from the rule. The goal of \tname\ is to create a
rule that only captures the negative traffic.

Figure~\ref{fig:overview} shows the workflow of \tname{} as a pipeline
of two components, appearing numbered in the figure. Each component
implements one of the steps mentioned above. The goal of the first
component is to produce a rule that captures the negative traffic.
First, it initializes the rule with random values. Only options
associated with the used protocol are included in the rule
representation. Then, \tname\ optimizes the rule options until
\suri\ is able to capture the attack associated with the negative
traffic.  \tname\ unsuccessfully terminates at this point if it cannot
capture the attack. The second component takes as input the rule
produced by the first component and tries to minimize that rule. Note
that any subset of options would capture the negative traffic---as all
options in the rule need to be satisfied for the negative traffic to
be captured---but it can also capture positive traffic. This component
systematically discards options from the rule encoding until no
positive traffic is captured.

\subsection{Positive Traffic}

Note from the figure that the negative traffic is an input to the
technique. The positive traffic, in contrast, is not. \tname{} uses
public databases storing different kinds of positive traffic for all
protocols covered by Suricata. \Fix{what databases?}
